### 2019-10-02 english-activity.Rmd
(jusque 29/09)

---
title: "My english activity log"
---

24/09/2019 : I read some articles for my presentation.

25/09/2019 :I read some articles for my presentation.

27/09/2019 :I read some articles for my presentation and I searched for some Youtube channels about science (especially Biology). I found : iBiology, Biomentors Classes Online, Beverly Biology, Vincent Racaniello, Alex Lee, Sea&amp;me - Marine Biology stuff with Maria.

29/09/2019 : I found other Youtube channels : Bleier Biology, Biology Bugbears, Bozeman science, Amoeba sisters, Free science lessons, Biomedical and Biological sciences.

From now on, you will use this article to include every single activity that you do in English in a chronological order (e.g. reading a book in English, watching TV shows and series in English, going to an English Evening in Mons, travelling to an English-speaking country, playing games in English such as the pretty exciting game LyricsTraining, etc.). You can even write a few lines about what you’ve watched/read/done and recommend certain things to your friends. The sky’s the limit! We value personal initiative and we will be happy to read whatever you have to offer.