### 2019-10-06 english-activity.Rmd
jusque 06/10

---
title: "My english activity log"
---

24/09/2019 : I read some articles for my presentation.

25/09/2019 :I read some articles for my presentation.

27/09/2019 :I read some articles for my presentation and I searched for some Youtube channels about science (especially Biology). I found : iBiology, Biomentors Classes Online, Beverly Biology, Vincent Racaniello, Alex Lee, Sea&amp;me - Marine Biology stuff with Maria.

29/09/2019 : I found other Youtube channels : Bleier Biology, Biology Bugbears, Bozeman science, Amoeba sisters, Free science lessons, Biomedical and Biological sciences.
I read the first chapter of "The Hobbit"

02/10/2019 : I read the wikipedia article about cockatiel for my Data Science homework.

05/10/2019 : I read some articles for my presentation and I watched a video about diffusion (https://www.youtube.com/watch?v=jhszFBtBPoI)

06/10/2019 : I read some articles for my presentation, I read the second chapter of "The Hobbit" and I watched a vide about cell cycle (https://www.youtube.com/watch?v=QVCjdNxJreE)