### 2019-10-07 _posts/2019-09-18-article-1/article-1.Rmd
éléphant retiré

---
title: "Article 1 : "
author:
  - name: Aurélie Pinchart
date: 09-18-2019
categories:
  - 
  - 
preview: image/
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

![]

## Vocabulary

| Words from the text | Synonym/explanation in English |
| ------------------- | ------------------------------ |
| E.g. To cope with   | To deal with                   |
|                     | Not too distant future         |

## Analysis table

|                              |                                                                     |
| ---------------------------- | ------------------------------------------------------------------- |
| Researchers                  | (insert names and, if possible, professions)                        |
| Published in?                | (insert date)                                                       |
| General topic?               |  (summarize the topic in one or two sentences)                      |
| Procedure/what was examined? | (Give some details about the topic, i.e. methodology/material/etc.) |
| Conclusions/discovery?       |  (what did the researchers conclude? Did they find something?)      |
| Remaining questions?         | (check if there are still interrogations at the end of the article) |