### 2019-10-20 _posts/2019-10-18-article-2/article-2.Rmd
jusqu'au 20/10/2019

---
title: "article-2"
description: |
  A short description of the post.
author:
  - name: Aurélie Pinchart
    url: {}
date: 10-18-2019
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill at <https://rstudio.github.io/distill>.