### 2019-10-21 _posts/2019-09-18-article-1/article-1.Rmd
suppression de l'article dans le texte

---
title: "Article 1 : Pathway found for treatment-resistant lung cancer"
author:
  - name: Aurélie Pinchart
date: 10-11-2019
categories:
  - track
preview: 
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Source of the article: <WWW.SCIENCEDAILY.COM/RELEASES/2019/09/190923170717.HTM>.

## Vocabulary

I used https://www.wordreference.com/, https://www.lexico.com/ and https://dictionary.cambridge.org to fill in this table.

| Vocabulary                      | Pronunciation     | Definition / synonym                                                               | French translation             |
| ------------------------------- | ----------------- | ---------------------------------------------------------------------------------- | ------------------------------ |
| The chemotherapy                | [ˌkiːməʊˈθɛrəpi]  | drug treatment of cancer                                                           | chimiothérapie                 |
| To prompt sb. / sth. to do sth. | [ˈprɒmpt]         | to encourage to do sht.                                                            | pousser qqn /qqch à faire qqch |
| The prognosis                   | [prɒgˈnəʊsɪs]     | doctor's judgement of your disease                                                 | pronostic                      |
| The immune system               | [ɪˈmjuːn ˈsɪstəm] | biological defense mechanisms                                                      | système immunitaire            |
| The mainstay                    | [ˈmeɪnsteɪ]       | most important part of something, providingsupport for everything else             | un pilier, un élément de base  |
| To recur                        | [rɪˈk3ːr]         | happen again                                                                       | se reproduire, réapparaître    |
| To enable                       | [ɪnˈeɪbəl]        | make possible                                                                      | permettre                      |
| Uncanny                         | [ˌʌnˈkæni]        | strange                                                                            | étrange, troublant             |
| To spit out (chemotherapy)      | [ˈspɪt ˈaʊt]      | reject the treatment                                                               | rejeter, résister à            |
| Dysfunctional                   |  [dɪsˈfʌŋkʃənəl]  | that does not work good                                                            | dysfonctionnel, anormal        |
| The chair                       | [ˈtʃɛər]          | person in charge if of meeting, official group or organization                     | le président (d'un comité)     |
| To target                       | [ˈtɑːrgɪt]        | aim an attack                                                                      | cibler                         |
|  frontline                      | [ˌfrʌntˈlaɪn]     | at the battlefront                                                                 | en première ligne              |
| The apoptosis                   | [ˌæpəpˈtoʊsəs]    | natural cell death                                                                 | apoptose                       |
| To knock sth. down              |  [ˈnɒk ˈdaʊn]     | destroy                                                                            | détruire                       |
| The antibody                    | [ˈæntibɒdi]       | blood protein that recognize pathogens                                             | anticoprs                      |
| To neutralize                   | [ˈnjuːtrəlaɪz]    | produce an effect that removes the effect of something else                        | neutraliser                    |
| To ensure sth.                  | [ɪnˈʃʊər]         | make sure that                                                                     | s'assurer que                  |
| Downstream                      | [ˌdaʊnˈstriːm]    |  happening later in a process or series of events                                  | en aval                        |
|  involved in                    | [ɪnˈvɒlvd ˈɪn]    | taking part in an activity                                                         | impliqué dans                  |
| The synergy                     | [ˈsɪnərdʒi]       | combined effect                                                                    | synergie                       |
| The outcome                     | [ˈaʊtkʌm]         | result                                                                             | résultat                       |
| Feedforward                     | [ˈfiːdfɔːwəd]     | predicting                                                                         | prédictif                      |
| Heavy-handed                    | [ˌhɛviˈhændɪd]    | severe                                                                             | brutal, sévère                 |
| To secrete                      | [sɪˈkriːt]        | emit                                                                               | sécréter                       |
| Adjacent                        | [əˈdʒeɪsənt]      | that is beside                                                                     | adjacent                       |
| The tissue                      |  [ˈtɪsjuː]        | a group of connected cells  that are similar to each other, have the same purpose  | tissu                          |
| A myriad of                     | [ˈmɪriəd]         | a large number                                                                     | une myriade de                 |

## Analysis table

|                             |                                                                                                                                                                                                                                                                                                                                                                           |
| --------------------------- | ------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------- |
| Researchers                 | Medical College of Georgia at Augusta University (lead authors : Dr. Wei Xiao, Dr. Mumtaz Rojiani, Dr. Amyn Rojiani)                                                                                                                                                                                                                                                      |
| Published in                | ScienceDaily, 23 September 2019                                                                                                                                                                                                                                                                                                                                           |
| General topic               | The way cancerous cells avoid apoptosis (non-small cell lung cancer)                                                                                                                                                                                                                                                                                                      |
| Procedure/what was examined | goal was to evidence link between TIMP-1 and resistance to chemotherapy + they used two drugs to destroy TIMP-1, then also destroy it by antibodies + added TIMP-1 to see the effect, then added IL-6  + STAT3 : regulator of gene activity (cell growth,division,  movement, apoptosis)                                                                                  |
| Conclusions/discovery       | high rate of TIMP-1 increases rate of IL-6 which make cells avoid apoptosis + low rate of TIMP-1 => lower rate of IL-6 => more apoptosis of cancerous cells => higher survival rate for patients + adding IL-6 has no real effect on apoptosis + remaining cancerous cells are more resistant + high level of IL-6 => STAT3 moves to nucleus => helps to reduce apoptosis |
| Remaining questions         | Does STAT3 activate more TIMP-1 production ?                                                                                                                                                                                                                                                                                                                              |