### 2019-11-03 _posts/2019-11-02-article-3/article-3.Rmd
jusqu'au 03/11/2019

---
title: "article-3"
description: |
  A short description of the post.
author:
  - name: Aurélie Pinchart
    url: {}
date: 11-02-2019
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill at <https://rstudio.github.io/distill>.