### 2019-11-24 _posts/2019-11-04-my-english-activity-log/my-english-activity-log.Rmd
jusqu'au 24/11/2019

---
title: "My English activity log"

author:
  - name: Aurélie Pinchart
date: 11-24-2019
preview: english.jpg
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

24/09/2019 : I read some articles for my presentation.

25/09/2019 :I read some articles for my presentation.

27/09/2019 :I read some articles for my presentation and I searched for some Youtube channels about science (especially Biology). I found : iBiology, Biomentors Classes Online, Beverly Biology, Vincent Racaniello, Alex Lee, Sea&amp;me - Marine Biology stuff with Maria.

29/09/2019 : I found other Youtube channels : Bleier Biology, Biology Bugbears, Bozeman science, Amoeba sisters, Free science lessons, Biomedical and Biological sciences.
I read the first chapter of "The Hobbit".

02/10/2019 : I read the wikipedia article about cockatiel for my Data Science homework.

05/10/2019 : I read some articles for my presentation and I watched a video about diffusion (https://www.youtube.com/watch?v=jhszFBtBPoI).

06/10/2019 : I read some articles for my presentation, I read the second chapter of "The Hobbit" and I watched a vide about cell cycle (https://www.youtube.com/watch?v=QVCjdNxJreE).

09/10/2019 : I read an article for my presentation.

10/10/2019 : I watched a video about the respiratory system (https://www.youtube.com/watch?v=2guHOnEt0e8).

11/10/2019 : I worked on my first article for my portfolio.

12/10/2019 : I read some articles for my presentation.

13/10/2019 : I worked on my presentation and I read the third chapter of "The Hobbit".

14/10/2019 : I read some articles for my presentation.

15/10/2019 : I read some articles for my presentation and I watched a documentary about coral bleaching https://www.youtube.com/watch?v=5cr8S5tTiz4.

16/10/2019 : I read some articles for my presentation.

17/10/2019 : I watched a video about the types of natural selection https://www.youtube.com/watch?v=DBPGLythma0.

19/10/2019 : I watched a video about marine biology https://www.youtube.com/watch?v=uWZXC3mjeZU. I also read the fouth chapter of "The Hobbit".

20/10/2019 : I watched a video about resting potential but it was said the opposite of what we were taught in class (so I don't put the link).

22/10/2019 : I watched a video about predation https://www.youtube.com/watch?v=6jXwEOit5vM.

24/10/2019 : I watched a video about genetic drift https://www.youtube.com/watch?v=W0TM4LQmoZY and a video about natural selection https://www.youtube.com/watch?v=7VM9YxmULuo.

26/10/2019 : I watched a video about the loss of biodiversity https://www.youtube.com/watch?v=vCkDxD0DV0Q and I read the fifth chapter of "T he Hobbit".

27/10/2019 : I watched a video about the protein structure and folding https://www.youtube.com/watch?v=hok2hyED9go.

29/10/2019 : I watched a video about mutations https://www.youtube.com/watch?v=vl6Vlf2thvI.

01/11/2019 : I watched a video about microscope drawings https://www.youtube.com/watch?v=NLKZ95aHqE4.

02/11/2019 : I watched a video about ethology https://www.youtube.com/watch?v=BvLihzeQzGk.

03/11/2019 : I watched a video about stress management https://www.youtube.com/watch?v=2vhIGYKW4gY and I read the sixth chapter of "The Hobbit".

05/11/2019 : I watched a video about ethology (the second part) https://www.youtube.com/watch?v=jEZIUVU0BCU.

07/11/2019 : I watched  the last part of the video about ethology https://www.youtube.com/watch?v=bspxe_-ilC4.

09/11/2019 : I watched a video about enzymes https://www.youtube.com/watch?v=6Acx6b3Hzsk 
10/11/2019 : I watched a video about germination https://www.youtube.com/watch?v=cmzhvl5kOc0 and I read the seventh chapter of "The Hobbit".

13/11/2019 : I watched a video about CRISPR https://www.youtube.com/watch?v=MnYppmstxIs.

15/11/2019 : I watched a video about ATP https://www.youtube.com/watch?v=23ZzI6WZS28.

16/11/2019 : I watched a video about bacteria https://www.youtube.com/watch?v=ORB866QSGv8.

17/11/2019 : I watched a video about autotrophs and heterotrophs https://www.youtube.com/watch?v=f8G7IulYxiA and I read the eighth of "The Hobbit".

20/11/2019 : I watched a video about DNA and RNA https://www.youtube.com/watch?v=JQByjprj_mA.

21/11/2019 : I watched a video about skin https://www.youtube.com/watch?v=b9Fim9kKVLs.

23/11/2019 : I watched a video about reflex arc https://www.youtube.com/watch?v=SMcF01utk3Y.

24/11/2019 : I watched a video about signal transduction pathways https://www.youtube.com/watch?v=236d5LDUFSs and I read the ninth chapter of "The Hobbit".