### 2019-12-14 _posts/2019-12-14-article-5/article-5.Rmd
article 5

---
title: "Article 5 : Possible new treatment strategy for lung cancer"
author:
  - name: Aurélie Pinchart
date: 12-14-2019
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Source of the article

H. Lee Moffitt Cancer Center & Research Institute. "Possible new treatment strategy for lung cancer: Combination therapy targeting KRAS mutation could provide benefit to nearly a third of lung cancer patients." ScienceDaily. ScienceDaily, 19 August 2019. <www.sciencedaily.com/releases/2019/08/190819140213.htm>.

Word count : 606

## Vocabulary

I used https://www.lexico.com/, https://dictionary.cambridge.org/ and https://www.wordreference.com/ to fill in this table.

| Vocabulary    | Pronunciation     | Definition / synonym                        | French translation           |
| ------------- | ----------------- | ------------------------------------------- | ---------------------------- |
| provide       | [prəˈvaɪd]        | give                                        | fournir                      |
| approach      | [əˈprəʊtʃ]        | the way of doing sth                        | approche                     |
| regulate      | [ˈrɛgjʊleɪt]      | control                                     | réguler                      |
| manner        | [ˈmænər]          | way, mode                                   | manière, façon               |
| occur         | [əˈk3ːr]          | happen                                      | se produire                  |
| melanoma      | [ˌmɛləˈnəʊmə]     | skin cancer                                 | mélanome                     |
| trametinib    | ?                 | cancer drug that inhibits MEK               | /                            |
| substantially | [səbˈstænʃəli]    | to a large extent                           | considérablement, énormément |
| dampen        | [ˈdæmpən]         | make feelings less strong (e.g. enthusiasm) | saper                        |
| sustain       | [səˈsteɪn]        | continue at the same level                  | soutenir, entreenir          |
| crucial       | [ˈkruːʃəl]        | very important                              | crucial                      |
| enhance       | [ɪnˈhɑːns]        | improve                                     | améliorer                    |
| viable        | [ˈvaɪəbəl]        | that can survive                            | viable                       |
| effectively   | [ɪˈfɛktɪvli]      | successfully                                | efficacement                 |
| conceptually  | [kənˈseptjʊəlɪ]   | based on abstract idea                      | conceptuellement             |
| novel         | [ˈnɒvəl]          | new                                         | nouveau, novateur            |
| oncogenic     | [ˌɒŋkə(ʊ)ˈdʒɛnɪk] | that causes the development of a tumor      | oncogénique                  |

## Analysis

|                               |                                                                                                                                                                                                                                                                                    |
| ----------------------------- | ---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------- |
| Researchers                   | Amer A. Beg and Mengyu Xie (Moffitt Cancer Center )                                                                                                                                                                                                                                |
| General topic                 | mutations in the KRAS pathway <-> 30% lung cancers                                                                                                                                                                                                                                 |
| Published in                  | Science Daily, 19 August 2019                                                                                                                                                                                                                                                      |
| Procedure / what was examined | MEK kinase = possible target for treatment  : drug screens + cell line + animal model experiments => MEK inhibitor trametinib combinated to cytokines TNFα and IFNγ => more cell death than MEK inhibitor alone                                                                    |
| Conclusion / discovery        | MEK inhibitors increase the number of receptors for TNFα +IFNγ improves the anti-tumor activity of MEK inhibitors and TNFα +MEK inhibitors increase expression of TNFα and IFNγ target genes + PD-1 immune checkpoint inhibitor with MEK inhibitors => strongest tumor regression  |
| Remaining questions           | further studies : combining MEK inhibitors and other immune therapies that increase expression of TNFα and IFNγ => find out if  modulation of the immune microenvironment along with tumor signaling is a good strategy to cure lung cancer patients with KRAS mutations           |