### 2020-02-15 _posts/2020-02-14-article-6/article-6.Rmd
article 6 + jusqu'au 15/02/2020

---
title: "Article 6 : Protein decoy stymies lung cancer growth in mice"
author:
  - name: Aurélie Pinchart
date: 02-14-2020
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Source of the article

Stanford Medicine. "Protein decoy stymies lung cancer growth in mice." ScienceDaily. ScienceDaily, 7 November 2019. <www.sciencedaily.com/releases/2019/11/191107160700.htm>

Word count : 930

## Vocabulary

I used https://www.lexico.com/, https://dictionary.cambridge.org/ and https://www.wordreference.com/ to fill in this table.

| Vocabulary        | Pronunciation         | Definition / synonym                         | French translation     |
| ----------------- | --------------------- | -------------------------------------------- | ---------------------- |
| decoy             | [ˈdiː.kɔɪ]            | sth. that diverts                            | leurre                 |
| stymie            | [ˈstaɪmi]             | block                                        | contrecarrer           |
| untreatable       | [ʌnˈtriːtəb(ə)l]      | that cannot be cured                         | incurable              |
| otherwise         | [ˈʌðərwaɪz]           | or esle                                      | sinon, autrement       |
| runaway           | [ˈrʌnəweɪ]            | out of control                               | hors de contrôle       |
| ligand            | [ˈlɪɡənd]             | molecule / ion attached on another molecule  | ligand                 |
| hook up           | [ˈhʊk ˈʌp]            | link                                         | s'attacher, s'associer |
| interfered        | [ˌɪntərˈfɪər]         | spoil / prevent the progress of a process    | interférer avec        |
| seek out          | [ˈsiːk ˈaʊt]          | search                                       | chercher               |
| thwart            | [ˈθwɔːrt]             | block                                        | contrecarrer           |
| interdisciplinary | [ˌɪntərˌdɪsɪˈplɪnəri] | multiple fields                              | interdisciplinaire     |
| squirting         | [ˈskw3ːrt]            | cause spurt of a liquid                      | asperger               |
| epithelial        | [ˌep.ɪˈθiː.li.əl]     | relating to epithelium                       | epithélial             |
| line              | [laɪn]                | cover                                        | recouvrir              |
| nested            | [nes′tid]             | one inside another                           | imbriqué, emboîté      |
| recruit           | [rɪˈkruːt]            | enroll                                       | recruter               |
| dampen            | ['dæmpən]             | make less strong / successful                | désenflammer           |
| unintentionally   | [ˌʌnɪnˈtɛnʃənəli]     | not deliberately                             | involontairement       |
| tightly           | [taɪtli]              | firmly                                       | fermement, solidement  |
| edge out          | [ɛdʒ ˈaʊt]            | displace, remove                             | évincer, déloger       |
| occur             | [əˈk3ːr]              | exist                                        | exister                |
| repel             | [rɪˈpɛl]              | keep away                                    | repousser              |
| painstaking       | ['peɪnzteɪkɪŋ]        | rigorous                                     | minutieux, rigoureux   |
| subtle            | ['sʌtəl]              | nearly undetectable                          | subtil                 |
| tweak             | [twiːk]               | slight change                                | petite modification    |
| propensity        | [prəʊˈpɛnsɪti]        | tendency                                     | propension, tendance   |
| ward off          | [ˈwɔːd ˈɒf]           | keep away                                    | repousser, éloigner    |
| instigate         | ['ɪnstɪgeɪt]          | cause                                        | être l'instigateur de  |
| prowess           | ['praʊəs/]            | great skill                                  | aptitude               |

## Analysis

|                               |                                                                                                                                                                                                         |
| ----------------------------- | ------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------- |
| Researchers                   | Jennifer Cochran, Alejandro Sweet-Cordero, Jun Kim and Cesar Marquez (Stanford and UC-San Francisco)                                                                                                    |
| General topic                 | Creating a decoy protein to stop a chain reaction (leading to the spread of cancer)                                                                                                                     |
| Published in                  | Science Daily, 7 November 2019                                                                                                                                                                          |
| Procedure / what was examined | In 1 tpye of non-small cell lung cancer : mutation in KRAS gene -> CLCF1 binds to 3 receptors (+activation) and is released by fibroblasts (induction by cancerous epithelial cells) -> mechanism ?     |
| Conclusion / discovery        | CLCF1 first binds to CNTFR -> this complex attracts the other receptors (gp130 and LIFR) => more energy for the tumor -> making a decoy receptor similar to CNTFR (but inactive) -> tested on mice : OK |
| Remaining questions           | Will it work on humans ?                                                                                                                                                                                |