### 2020-02-29 _posts/2020-02-29-article-7/article-7.Rmd
article 7

---
title: "Article 7 : Newly developed nanoparticles help fight lung cancer in animal model"
author:
  - name: Aurélie Pinchart
date: 02-29-2020
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Source of the article

Source : Wake Forest Baptist Medical Center. "Newly developed nanoparticles help fight lung cancer in animal model." ScienceDaily. ScienceDaily, 14 November 2019. <www.sciencedaily.com/releases/2019/11/191114115938.htm>.

Word count : 535

## Vocabulary
I used https://www.lexico.com/, https://dictionary.cambridge.org/, https://www.wordreference.com/, accentperfect.com/ to fill in this table

| Vocabulary        | Pronunciation       | Definition / synonym            | French translation          |
| ----------------- | ------------------- | ------------------------------- | --------------------------- |
| inhale            | [ɪnˈheɪl]           | breathe in                      | inspirer                    |
| metastatic        | [ˌmɛtəˈstætɪk]      | spreading                       | métastatique                |
| respond (to)      | [rɪˈspɒnd]          | react                           | réagir (à)                  |
| inhalable         | [ɪ’nheɪləbəl]       | that can be inhale              | inhalable                   |
| infiltrated       | [ˈɪnfɪltreɪt]       | move into sth.                  | s'infiltrer dans            |
| immunosuppressive | [ˌɪmjʊnəʊsəˈprɛsɪv] | drug that weakens immune system | immunosuppresseur           |
| novel             | [ˈnɒvəl]            | new, different                  | nouveau, novateur           |
| convert (to)      | [kənˈv3ːrt]         | change (to)                     | transformer, convertir (en) |
| deposit           | [dɪˈpɒzɪt]          | place                           | déposer                     |
| (lung) air sac    | [ˈeə ˌsæk]          | alveolus                        | alvéole pulmonaire          |
| systemic          | [sɪˈstɛmɪk]         | affecting all the body          | systémique, généralisé      |
| regression        | [rɪˈgrɛʃən]         | reverting to a previous state   | régression                  |
| prolong           | [prəʊˈlɒŋ]          | make longer                     | prolonger                   |
| eliminate         | [ɪˈlɪmɪneɪt]        | get rid of                      | éliminer, faire disparaître |
| robust            | [ˈrəʊbʌst]          | strong                          | robuste                     |
| deep-seated       | [ˌdiːpˈsiːtɪd]      | at a deep level                 | fermement ancré, profond    |
| feasibility       | [ˌfiːzəˈbɪlɪti]     | possibility of doing            | faisabilité                 |
| provisional       | [prəˈvɪʒənəl]       | temporary                       | provisoire                  |
| patent            | [ˈpeɪtənt]          | legal right                     | brevet                      |

## Analysis

|                               |                                                                                                                                                                                                                                                       |
| ----------------------------- | ----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------- |
| Researchers                   | Dawen Zhao (Wake Forest School of Medicine)                                                                                                                                                                                                           |
| General topic                 | Inhalable nano-particles to treat lung cancer                                                                                                                                                                                                         |
| Published in                  | Science Daily, 14 November 2019                                                                                                                                                                                                                       |
| Procedure / what was examined | mice, inhalable particles (immunotherapy; non-invasive, access not only to the surface) -> taken up by APC cells => activation of immune response + radiation therapy; injection of immunomodulators not so efficient (<-> only surface of the tumor) |
| Conclusion / discovery        | regression of tumors + longer survival; elimination of tumors in some cases                                                                                                                                                                           |
| Remaining questions           | /                                                                                                                                                                                                                                                     |