### 2020-04-18 _posts/2020-04-18-introduction/introduction.Rmd

---
title: "Introduction"
author:
  - name: Aurélie Pinchart
date: 04-18-2020
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

There are two main types of lung cancer: small-cell lung cancer and non-small-cell lung cancer. The symptoms are coughing, coughing up blood, weight loss, shortness of breath and chest pain. About 85% of lung cancers are due to tobacco smoking and the 15% left impact people who have never smoked and is due to genetic factors or pollution (such as second-hand smoke).

This disease can be diagnosed thanks to radiographs, computed tomography scan and biopsy. Lung cancer is a great cause of cancer related-death. 
My own grandfather died a few years ago because of it – he used to smoke a lot when I was a child, then he stopped but it seems I was too late… So I wanted to know a bit more about this disease. I especially wanted to know more about new discoveries in molecular and cellular biology to get an idea of how a lung cancer cell works and then also get an idea of future possible treatments. I also wanted to know if screening techniques had improved because waiting for many symptoms to diagnose lung cancer is sometimes too late to treat it properly.

In this portfolio, you will find 8 articles about recent researches mostly on non-small-cell lung cancer explaining how these cells work, what the possible future treatments are and how we can detect if you have lung cancer.